---
title: "Comparison of different copula structures on mulitquestion simulated data from Irish registry and PROACT"
format: html
editor: source
---

```{r setup}
#| include: FALSE
library(tidyverse)
library(knitr)
library(ggdist)
library(patchwork)

source(here::here("R", "post_epred_hetsked.R"))

cache_dir <- here::here("local_temp_data", "simulation_test")


# manually set these to match fits
M <- 400
eff_probs <- rep(c(0.5, 0.75), times = M)
n_subjects_per_group <- 50

# load data
irlsim_ls <- readRDS(paste0(cache_dir, "/sim_irldata_4Q.RDS"))
proact_ls <- readRDS(paste0(cache_dir, "/sim_proactdata_4Q.RDS"))


irl_maxs <- irlsim_ls |>
    map(\(dat)
        dat |>
            group_by(subject_id) |>
            filter(visit_id == max(visit_id)) |>
            ungroup() |>
            select(visit_id))

pro_maxs <- proact_ls |>
    map(\(dat)
        dat |>
            group_by(subject_id) |>
            filter(visit_id == max(visit_id)) |>
            ungroup() |>
            select(visit_id))

```

## Data summary

There were `r length(irlsim_ls)` datasets simulated from the Irish register, and `r length(proact_ls)` datasets simulated from the PROACT data with `r length(unique(eff_probs))` distinct effective sizes used. The number of responses per subject over time are summarised in the following plot:

```{r repliesplot}
#| echo: FALSE

plotdat <- bind_rows(irl_maxs |>
              bind_cols(.name_repair = c("unique_quiet")) |>
              mutate(source = "irish"),
          pro_maxs |>
              bind_cols(.name_repair = c("unique_quiet")) |>
              mutate(source = "proact")) |>
    pivot_longer(cols = -source,
                 names_prefix = "visit_id...",
                 names_to = "m")

p1 <- ggplot(plotdat, aes(x = value)) +
    geom_histogram(bins = max(plotdat$value)) +
    ggtitle("Number of longitudinal timepoints") + 
    theme_minimal() +
    facet_wrap(~source, ncol = 1)
print(p1)

```


```{r loadpredsummarise}
#| include: FALSE

# define new_data for fit
new_data <- data.frame(time2 = c(0, 12, 0, 12),
                       group = c(0, 0, 1, 1)) |>
    mutate(interaction = group * time2)


# Load fits and predict effect sizes
irlfits1 <- readRDS(paste0(cache_dir, "/", "2026-08-31_compare_copula_Q4_n50_mvprob_400.RDS"))
effects_irl1 <- irlfits1 |>
    map(\(fit) {
        epreds <- post_epred_mvprobit_hetsked(fit, new_data, mode = "homoskedastic")
        sum_epreds <- apply(epreds, c(1, 2), sum) -12
        (sum_epreds[,4] - sum_epreds[,2]) - (sum_epreds[,3] - sum_epreds[,1])
    })
rm(irlfits1) ; invisible(gc())

irlfits2 <- readRDS(paste0(cache_dir, "/", "2026-08-31_compare_copula_Q4_n50_mvprobhet_400.RDS"))
effects_irl2 <- irlfits2 |>
    map(\(fit) {
        epreds <- post_epred_mvprobit_hetsked(fit, new_data, mode = c("heteroskedastic"))
        sum_epreds <- apply(epreds, c(1, 2), sum) -12
        (sum_epreds[,4] - sum_epreds[,2]) - (sum_epreds[,3] - sum_epreds[,1])
    })
rm(irlfits2) ; invisible(gc())

irlfits3 <- readRDS(paste0(cache_dir, "/", "2026-08-31_compare_copula_Q4_n50_mvstudt_n3_400.RDS"))
effects_irl3 <- irlfits3 |>
    map(\(fit) {
        epreds <- post_epred_studthetsked(fit, new_data, nu =3, mode = c("homoskedastic"))
        sum_epreds <- apply(epreds, c(1, 2), sum) -12
        (sum_epreds[,4] - sum_epreds[,2]) - (sum_epreds[,3] - sum_epreds[,1])
    })
rm(irlfits3) ; invisible(gc())

irlfits4 <- readRDS(paste0(cache_dir, "/", "2026-08-31_compare_copula_Q4_n50_mvstudt_nubeta0.1_400.RDS"))
effects_irl4 <- irlfits4 |>
    map(\(fit) {
        epreds <- post_epred_studthetsked(fit, new_data, mode = c("homoskedastic"))
        sum_epreds <- apply(epreds, c(1, 2), sum) -12
        (sum_epreds[,4] - sum_epreds[,2]) - (sum_epreds[,3] - sum_epreds[,1])
    })
rm(irlfits4) ; invisible(gc())

irlfits5 <- readRDS(paste0(cache_dir, "/", "2026-08-31_compare_copula_Q4_n50_mvstudt_nubeta0.9_400.RDS"))
effects_irl5 <- irlfits5 |>
    map(\(fit) {
        epreds <- post_epred_studthetsked(fit, new_data,mode = c("homoskedastic"))
        sum_epreds <- apply(epreds, c(1, 2), sum) -12
        (sum_epreds[,4] - sum_epreds[,2]) - (sum_epreds[,3] - sum_epreds[,1])
    })
rm(irlfits5) ; invisible(gc())

irlfits6 <- readRDS(paste0(cache_dir, "/", "2026-08-31_compare_copula_Q4_n50_mvstudthetsked_nu3_400.RDS"))
effects_irl6 <- irlfits6 |>
    map(\(fit) {
        epreds <- post_epred_studthetsked(fit, new_data, nu = 3.0, mode = c("het2par"))
        sum_epreds <- apply(epreds, c(1, 2), sum) -12
        (sum_epreds[,4] - sum_epreds[,2]) - (sum_epreds[,3] - sum_epreds[,1])
    })
rm(irlfits6) ; invisible(gc())



# load predict effects PROACT fit
proactfits1 <- readRDS(paste0(cache_dir, "/", "2026-09-01_PROACT_compare_copula_Q4_n50_mvprob_400.RDS"))
effects_proact1 <- proactfits1 |>
    map(\(fit) {
        epreds <- post_epred_mvprobit_hetsked(fit, new_data, mode = "homoskedastic")
        sum_epreds <- apply(epreds, c(1, 2), sum) -12
        (sum_epreds[,4] - sum_epreds[,2]) - (sum_epreds[,3] - sum_epreds[,1])
    })
rm(proactfits1) ; invisible(gc())

proactfits2 <- readRDS(paste0(cache_dir, "/", "2026-09-01_PROACT_compare_copula_Q4_n50_mvprobhet_400.RDS"))
effects_proact2 <- proactfits2 |>
    map(\(fit) {
        epreds <- post_epred_mvprobit_hetsked(fit, new_data, mode = c("heteroskedastic"))
        sum_epreds <- apply(epreds, c(1, 2), sum) -12
        (sum_epreds[,4] - sum_epreds[,2]) - (sum_epreds[,3] - sum_epreds[,1])
    })
rm(proactfits2) ; invisible(gc())

proactfits3 <- readRDS(paste0(cache_dir, "/", "2026-09-01_PROACT_compare_copula_Q4_n50_mvstudt_n3_400.RDS"))
effects_proact3 <- proactfits3 |>
    map(\(fit) {
        epreds <- post_epred_studthetsked(fit, new_data, nu =3, mode = c("homoskedastic"))
        sum_epreds <- apply(epreds, c(1, 2), sum) -12
        (sum_epreds[,4] - sum_epreds[,2]) - (sum_epreds[,3] - sum_epreds[,1])
    })
rm(proactfits3) ; invisible(gc())

proactfits4 <- readRDS(paste0(cache_dir, "/", "2026-09-01_PROACT_compare_copula_Q4_n50_mvstudt_nubeta0.1_400.RDS"))
effects_proact4 <- proactfits4 |>
    map(\(fit) {
        epreds <- post_epred_studthetsked(fit, new_data, mode = c("homoskedastic"))
        sum_epreds <- apply(epreds, c(1, 2), sum) -12
        (sum_epreds[,4] - sum_epreds[,2]) - (sum_epreds[,3] - sum_epreds[,1])
    })
rm(proactfits4) ; invisible(gc())

proactfits5 <- readRDS(paste0(cache_dir, "/", "2026-09-01_PROACT_compare_copula_Q4_n50_mvstudt_nubeta0.9_400.RDS"))
effects_proact5 <- proactfits5 |>
    map(\(fit) {
        epreds <- post_epred_studthetsked(fit, new_data,mode = c("homoskedastic"))
        sum_epreds <- apply(epreds, c(1, 2), sum) -12
        (sum_epreds[,4] - sum_epreds[,2]) - (sum_epreds[,3] - sum_epreds[,1])
    })
rm(proactfits5) ; invisible(gc())

proactfits6 <- readRDS(paste0(cache_dir, "/", "2026-09-01_PROACT_compare_copula_Q4_n50_mvstudthetsked_nu3_400.RDS"))
effects_proact6 <- proactfits6 |>
    map(\(fit) {
        epreds <- post_epred_studthetsked(fit, new_data, nu = 3.0, mode = c("het2par"))
        sum_epreds <- apply(epreds, c(1, 2), sum) -12
        (sum_epreds[,4] - sum_epreds[,2]) - (sum_epreds[,3] - sum_epreds[,1])
    })
rm(proactfits6) ; invisible(gc())


```


```{r summary}
#| include: FALSE

##### summarise predictions irl
summ_irl1 <- map_df(1:length(eff_probs),
                       ~ data.frame(id = .x, eff_prob = eff_probs[[.x]],
                                    estimate = effects_irl1[[.x]],
                                    family = "mvprobit", df = "",
                                    copula = "homosked", data = "irl"))
summ_irl2 <- map_df(1:length(eff_probs),
                       ~ data.frame(id = .x, eff_prob = eff_probs[[.x]],
                                    estimate = effects_irl2[[.x]],
                                    family = "mvprobit", df = "",
                                    copula = "heterosked", data = "irl"))
summ_irl3 <- map_df(1:length(eff_probs),
                       ~ data.frame(id = .x, eff_prob = eff_probs[[.x]],
                                    estimate = effects_irl3[[.x]],
                                    family = "mvstudt", df = "3",
                                    copula = "homosked", data = "irl"))
summ_irl4 <- map_df(1:length(eff_probs),
                       ~ data.frame(id = .x, eff_prob = eff_probs[[.x]],
                                    estimate = effects_irl4[[.x]],
                                    family = "mvstudt", df = "prior0.1",
                                    copula = "homosked", data = "irl"))
summ_irl5 <- map_df(1:length(eff_probs),
                       ~ data.frame(id = .x, 
                                    eff_prob = eff_probs[[.x]],
                                    estimate = effects_irl5[[.x]],
                                    family = "mvstudt", df = "prior0.9",
                                    copula = "homosked", data = "irl"))
summ_irl6 <- map_df(1:length(eff_probs),
                       ~ data.frame(id = .x,
                                    eff_prob = eff_probs[[.x]],
                                    estimate = effects_irl6[[.x]],
                                    family = "mvstudt", df = "3",
                                    copula = "heterosked", data = "irl"))

##### summarise predictions proact
summ_proact1 <- map_df(1:length(eff_probs),
                    ~ data.frame(id = .x, eff_prob = eff_probs[[.x]],
                                 estimate = effects_proact1[[.x]],
                                 family = "mvprobit", df = "",
                                 copula = "homosked", data = "proact"))
summ_proact2 <- map_df(1:length(eff_probs),
                    ~ data.frame(id = .x, eff_prob = eff_probs[[.x]],
                                 estimate = effects_proact2[[.x]],
                                 family = "mvprobit", df = "",
                                 copula = "heterosked", data = "proact"))
summ_proact3 <- map_df(1:length(eff_probs),
                    ~ data.frame(id = .x, eff_prob = eff_probs[[.x]],
                                 estimate = effects_proact3[[.x]],
                                 family = "mvstudt", df = "3",
                                 copula = "homosked", data = "proact"))
summ_proact4 <- map_df(1:length(eff_probs),
                    ~ data.frame(id = .x, eff_prob = eff_probs[[.x]],
                                 estimate = effects_proact4[[.x]],
                                 family = "mvstudt", df = "prior0.1",
                                 copula = "homosked", data = "proact"))
summ_proact5 <- map_df(1:length(eff_probs),
                    ~ data.frame(id = .x, 
                                 eff_prob = eff_probs[[.x]],
                                 estimate = effects_proact5[[.x]],
                                 family = "mvstudt", df = "prior0.9",
                                 copula = "homosked", data = "proact"))
summ_proact6 <- map_df(1:length(eff_probs),
                    ~ data.frame(id = .x,
                                 eff_prob = eff_probs[[.x]],
                                 estimate = effects_proact6[[.x]],
                                 family = "mvstudt", df = "3",
                                 copula = "heterosked", data = "proact"))

# combine summaries
all_summs <- rbind(summ_irl1, summ_irl2, summ_irl3,
                   summ_irl4, summ_irl5, summ_irl6,
                   summ_proact1, summ_proact2, summ_proact3,
                   summ_proact4, summ_proact5, summ_proact6)

# summaries the summaries :D
summs <- all_summs |>
    summarise(.by = c(id, data, family, copula, df),
              eff_prob = unique(eff_prob),
              q_025 = quantile(estimate, 0.025),
              mean = mean(estimate),
              q_975 = quantile(estimate, 0.975),
              flag1 = ifelse(eff_prob == 0.5 & q_025 > 0 | q_975 < 0, 1, 0),
              flag2 = ifelse(q_025 > 0, 1, 0)) |>
    summarise(.by = c(family, data, copula, df, eff_prob),
              type1error = mean(flag1, na.rm = TRUE),
              power = mean(flag2)) 

```


## Summary statistics of models

#### Type 1 error
```{r showsummaries}
#| echo: FALSE
summs |>
    filter(eff_prob == 0.5) |>
    dplyr::select(family, data, copula, df, eff_prob, type1error) |>
    pivot_wider(id_cols = c(family, copula, df, eff_prob),
                names_from = data, values_from = type1error,
                names_prefix = "T1err_") |>
    arrange(T1err_irl) |>
    kable()

```




#### Power
```{r}
#| echo: FALSE

summs |>
    filter(eff_prob == 0.75) |>
    dplyr::select(family, data, copula, df, eff_prob, power) |>
    pivot_wider(id_cols = c(family, copula, df, eff_prob),
                names_from = data, values_from = power,
                names_prefix = "Power_") |>
    arrange(Power_irl) |>
    kable()

```


#### Visualisation of recovered effect sizes

```{r plotT1err}
#| echo: FALSE
#| message: false

plot_T1err_irl_0.5 <- all_summs |>
    filter(eff_prob == 0.5 & data == "irl") |>
    mutate(eff_prob = factor(eff_prob),
           label = paste0(family, "\n", copula, " ", df)) |>
    ggplot(aes(x =  id, y = estimate, col = copula)) +
    geom_hline(yintercept = 0) +
    stat_pointinterval(position = position_dodge(width = 0.3)) +
    theme_minimal() +
    ggtitle("Simulated from Irish cohort") +
    facet_wrap(~ label, ncol = 6)#, scales = "free_y")

plot_T1err_proact_0.5 <- all_summs |>
    filter(eff_prob == 0.5 & data == "proact") |>
    mutate(eff_prob = factor(eff_prob),
           label = paste0(family, "\n", copula, " ", df)) |>
    ggplot(aes(x =  id, y = estimate, col = copula)) +
    geom_hline(yintercept = 0) +
    stat_pointinterval(position = position_dodge(width = 0.3)) +
    theme_minimal() +
    ggtitle("Simulated from PROACT data") +
    facet_wrap(~ label, ncol = 6)

plot_T1err_irl_0.75 <- all_summs |>
    filter(eff_prob == 0.75 & data == "irl") |>
    mutate(eff_prob = factor(eff_prob),
           label = paste0(family, "\n", copula, " ", df)) |>
    ggplot(aes(x =  id, y = estimate, col = copula)) +
    geom_hline(yintercept = 0) +
    stat_pointinterval(position = position_dodge(width = 0.3)) +
    theme_minimal() +
    ggtitle("Simulated from Irish cohort") +
    facet_wrap(~ label, ncol = 6)#, scales = "free_y")

plot_T1err_proact_0.75 <- all_summs |>
    filter(eff_prob == 0.75 & data == "proact") |>
    mutate(eff_prob = factor(eff_prob),
           label = paste0(family, "\n", copula, " ", df)) |>
    ggplot(aes(x =  id, y = estimate, col = copula)) +
    geom_hline(yintercept = 0) +
    stat_pointinterval(position = position_dodge(width = 0.3)) +
    theme_minimal() +
    ggtitle("Simulated from PROACT data") +
    facet_wrap(~ label, ncol = 6)

plot_eff0.5 <- plot_T1err_irl_0.5 + plot_T1err_proact_0.5 +
    plot_layout(guides = "collect", ncol = 1) +
    plot_annotation(title = "Difference in expected value by group",
                    subtitle = paste0(M, " simulations,",
                                      n_subjects_per_group,
                                      " subjects per group",
                                      ", simulated effect prob. = 0.5"))
print(plot_eff0.5)


plot_eff0.75 <- plot_T1err_irl_0.75 + plot_T1err_proact_0.75 +
    plot_layout(guides = "collect", ncol = 1) +
    plot_annotation(title = "Difference in expected value by group",
                    subtitle = paste0(M, " simulations,",
                                      n_subjects_per_group,
                                      " subjects per group",
                                      ", simulated effect prob. = 0.75"))
print(plot_eff0.75)




```

