---
title: "Examine_results 12Q"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tableone)
library(brms)
library(bayesplot)
library(rstan)
library(corrplot)
library(here)
library(kableExtra)

cache_dir <- here("local_temp_data")

# Load data
df_frs <- readRDS(here("Data/long_format_wide12Qs.RDS"))
frs_long <- readRDS(here("Data/long_format_long12Qs.RDS"))


# Load models
fit_mvprob12Q <- readRDS(paste0(cache_dir, "/mvprobit_bgoodri_12Qs.rds"))

# Load correlation helper functions
source(here("R/helpfns.R"))
source(here("R/pp_checks.R"))
source(here("R/mv_probit.R"))


questions <- c("Q01", "Q02", "Q03", "Q04", "Q05", "Q06",
               "Q07", "Q08", "Q09", "Q10", "Q11", "Q12")

# Get correlations using helper function
# need to modify this to alllow subsets of data
cor_bgood <- predcor_multv(fit_mvprob12Q, timevar = "alsfrs_dly_mnths")

```

The model had the following numbers of parameters:
```{r numpars, echo=FALSE}
print(paste0("Model 6 (mvpro_bgood): ", length(parnames(fit_mvprob12Q))))

```
and the following numbers of divergences:
```{r divs, echo=FALSE}
print(paste0("Model 6: ", get_num_divergent(fit_mvprob12Q$fit)))

```

lets look at a traceplot or 2:
```{r trace}

mcmc_trace(fit_mvprob12Q, pars = c("b_Q01_Intercept[3]", "b_Q07_alsfrs_dly_mnths",
                                   "cor_ID__Q06_Intercept__Q08_alsfrs_dly_mnths",
                                   "r_1[74,20]","Rescor[10,12]"))


```


The times for warmup and sampling for each chain of the models were:
```{r times, echo=FALSE}
print("Model 6:")
get_elapsed_time(fit_mvprob12Q$fit)
```



## Model 6 - Multivariate probit - bgoodri's parametrization
```{r mod6_bg, echo=FALSE}
kable( cor_bgood , digits = 2, table.attr = "style='width:75%;'") 
corrplot(cor_bgood , type = "lower")

pred_rescor_bgoodri <- posterior_predict_mv_probit(fit_mvprob12Q)

pp_check_cor_wide(pred_rescor_bgoodri, df_frs, questions, actual_point_size = 3) 
pp_check_cor_wide(pred_rescor_bgoodri, df_frs, questions, group = as.integer(df_frs$site == "Spinal"))


```

How many rows (individuals) in the data ?
```{r nobs}
nrow(df_frs)
```


Let's look at the fixed effects:
```{r fixef}
summ_fixef <- fixef(fit_mvprob12Q) %>% 
    data.frame()
summ_fixef$param <- row.names(summ_fixef)
row.names(summ_fixef) <- NULL

summ_fixef <- summ_fixef %>% 
    dplyr::select(param, Estimate, Est.Error, Q2.5, Q97.5)

#
# summarise the intercepts
summ_fixef %>%
    filter( grepl( "Intercept", summ_fixef$param, fixed = TRUE) ) %>% 
    kable(digits = 2)



#summarise the slopes
summ_fixef %>%
    filter( grepl( "alsfrs_dly_mnths", summ_fixef$param, fixed = TRUE) ) %>% 
    kable(digits = 2)

# summarise the age covar
summ_fixef %>%
    filter( grepl( "age_dx", summ_fixef$param, fixed = TRUE) ) %>% 
    kable(digits = 2)

# summarise the dx_delay covar
summ_fixef %>%
    filter( grepl( "dx_delay", summ_fixef$param, fixed = TRUE) ) %>% 
    kable(digits = 2)


```



How does longitudinal total score compare to raw total score ?
```{r ppc_long}

temp <- apply(pred_rescor_bgoodri, c(2, 3), mean) # can make this faster
temp2 <- rowSums(temp) - 12 # -12 because 1 added to every question
df_frs$fit <- temp2

ggplot(df_frs, aes(x=Total, y = fit)) + geom_point() +
    coord_fixed() + facet_wrap(~site) +
    geom_abline(slope=1, intercept=0)
```

How would this compare to an lmer model as would traditionally do ?
```{r}
library(lme4)
library(splines)
mm1 <- lmer(Total ~ ns(alsfrs_dly_mnths,2) + (ns(alsfrs_dly_mnths,2)|ID),
            df_frs)

df_frs$fit_lmer <- predict(mm1)

ggplot(df_frs, aes(x=Total, y = fit_lmer)) + geom_point() +
    coord_fixed() + facet_wrap(~site) +
    geom_abline(slope=1, intercept=0)                        
                           
```


## Compare fit over time
```{r time_byfit}
# raw data
ggplot(df_frs, aes(x = alsfrs_dly_mnths, y = Total, group=ID)) + geom_line() +
    facet_wrap(~site)

# raw data + fit
ggplot(df_frs, aes(x = alsfrs_dly_mnths, y = Total, group=ID)) + geom_line() +
    geom_line(aes(y = fit, col="mv_probit")) +
    geom_line(aes(y = fit_lmer, col="lme4")) +
    facet_wrap(~site)


```



